---
title: "day-9-exercise"
author: "Alex Smilor"
format: 
  html:
    self-contained: true
execute:
  echo: true
---
```{r}
knitr::opts_chunk$set(echo = TRUE)
```

The Ozone data probably doesn't require cleaning, though it has some missing values. In this case, I think it would be best to simply drop the NA values
```{r}
#Setup
library(tidyverse)
library(visdat)
library(broom)
library(ggthemes)
#In the future, remember not to include install.packages in a qmd file, it breaks the rendering

airquality <- airquality
vis_dat(airquality)
#The Ozone data probably doesn't require cleaning, though it has some missing values. In this case, I think it would be best to simply drop the NA values

#Removal
airquality <- airquality %>% 
  drop_na()
```

I chose to model the amount of Ozone in the atmosphere as a function of daily high temperature, since warmer days increase the rate of the chemical reactions that cause ozone formation. The below model appears to valid and shows a roughly linear trend in the data. The R^2 = 0.48, meaning that approximately 48% of the variation in the mean amount of ozone is explained by the maximum daily temperature. 
```{r}
#Linear regression model of Ozone predicted by the maximum daily temperature
model <- airquality %>% 
  lm(formula = Ozone ~ Temp)
summary(model) #Summary statistics; R^2 = 0.48

augment <- augment(model, airquality)
#Plot of Actual vs Predicted with a correlation of 0.7
ggplot(augment)+
  geom_point(aes(x=Ozone, y = .fitted)) +
  geom_abline(intercept = 0, slope = 1, color = "red")+ 
  labs(subtitle = paste("Correlation:", round(cor(augment$Ozone, augment$.fitted),2)),
       x = "Actual",
       y = "Predicted")+
  ggthemes::theme_clean()
```

